---
title: "IST 707 Project"
author: "Alex Klein & Marley Akonnor"
date: "11/18/2021"
output:
  word_document: default
  pdf_document: default
---

#INSTALL PACKAGES USED
```{r}
library("gdata")
library("boot")
library("plyr")
library("readxl")
library("randomForest")
library("ggplot2")
library("pROC")
library("reshape2")
library("dplyr")
library("forecast")
library("neuralnet")
library("caret")
library("Metrics")
```

#IMPORT DATASET AND SAVE AS DATAFRAME
```{r}
DataSetDataFrame1 <- read_excel("C:/Users/klein/Desktop/SU Classes/Quarter 2/IST 707/Project/cwurData.xlsx")
DF1 <- data.frame(DataSetDataFrame1)
```

#CLEAN AND OBSERVE DATA
```{r}
DF1 <- na.omit(DF1)
str(DF1)
```

#REMOVE THE BELOW 4 COLUMNS FOR EDA
```{r}
DF1_EDA <- DF1[, !(names(DF1) %in% c("institution", "country", "broad_impact", "year"))]
```

#CREATE A CORRELATION MATRIX
```{r}
pairs(DF1_EDA[,3:10])
```

#CREATE RANDOMFOREST
#CREATE X AND Y MEASURES FOR RANDOM FOREST
```{r}
XF <- DF1 %>%
  select(quality_of_education, alumni_employment, quality_of_faculty, publications, influence, citations, broad_impact, patents)
YT <- DF1$score
```

#CREATE TRAINING AND TESTING DATASETS
```{r}
INDEXRF <- createDataPartition(YT, p=0.6, list=FALSE)
XF_TRAINING <- XF[INDEXRF,]
XF_TESTING <- XF[-INDEXRF,]
YT_TRAINING <- YT[INDEXRF]
YT_TESTING <- YT[-INDEXRF]
```

#TRAINING THE RANDOMFOREST
```{r}
RFOUTPUT <- randomForest(x=XF_TRAINING, y=YT_TRAINING, maxnodes=10,ntree=10)
```

#CREATING PREDICTIONSRF
```{r}
PREDICTIONSRF <- predict(RFOUTPUT, XF_TESTING)

RESULTSRF <- XF_TESTING
RESULTSRF['score'] <- YT_TESTING
RESULTSRF['prediction']<-  PREDICTIONSRF

head(RESULTSRF)
```


#CREATING SCATTERPLOT GRAPH
```{r}
ggplot(  ) + 
  geom_point( aes(x = XF_TESTING$patents, y = YT_TESTING, color = 'red', alpha = 0.5) ) + 
  geom_point( aes(x = XF_TESTING$patents , y = PREDICTIONSRF, color = 'blue',  alpha = 0.5)) + 
  labs(x = "patents", y = "Score", color = "", alpha = 'Transperency') +
  scale_color_manual(labels = c( "Predicted", "Real"), values = c("blue", "red")) 
```


#FINDING MEAN SQUARED ERROR, and R-SQUARED VALUES

```{r}
print(paste0('RMSE: ' ,caret::postResample(PREDICTIONSRF , YT_TESTING)['RMSE']^2 ))
```

```{r}
print(paste0('R2: ' ,caret::postResample(PREDICTIONSRF , YT_TESTING)['Rsquared'] ))
```

# Normalized RMSE
```{r}
# NRMSE = RMSE / ( max value – min value )
NRMSE = 2.39147488085974 / (100.00 - 44.02)
NRMSE
```

#Create DF2 w No Non-Numeric fields.
```{r}
DF2 <- DF1[,c(-1:-4,-14)]
```

#CREATING SAMPLES, TRAINING, AND TESTING DATASETS FOR NEURAL NETWORK
```{r}
samplesize = 0.60 * nrow(DF2)
set.seed(76)
INDEXNN = sample( seq_len ( nrow ( DF2 ) ), size = samplesize )

TRAINDFNN = DF2[ INDEXNN, ]
TESTDFNN = DF2[ -INDEXNN, ]
```

#SCALING/NORMALIZING DATA FOR NEURAL NETWORK
```{r}
MAXIMUM = apply(DF2 , 2 , max)
MINIMUM = apply(DF2, 2 , min)
scaled = as.data.frame(scale(DF2, center = MINIMUM, scale = MAXIMUM - MINIMUM))
```

#CREATING DATASETS FOR TESTING AND TRAINING NEURAL NETWORK
```{r}
NNTRAINING = scaled[INDEXNN , ]
NNTESTING = scaled[-INDEXNN , ]

set.seed(7)
NN = neuralnet(score~quality_of_education + alumni_employment+ quality_of_faculty+ publications+ influence+ citations+ broad_impact+ patents, NNTRAINING, hidden = c(6,4,2), linear.output = FALSE )

plot(NN)
```

#CREATING PREDICTIONS FROM NEURAL NETWORK
```{r}
PRESULTS <- subset(NNTESTING, select = c("quality_of_education", "alumni_employment", "quality_of_faculty", "publications", "influence", "citations", "broad_impact", "patents"))
head(PRESULTS)
NNPRESULTS <- compute(NN, PRESULTS)
NNRESULTS <- data.frame(actual = NNTESTING$score, prediction = NNPRESULTS$net.result)
head(NNRESULTS)
```

#GRAPHING PREDICTIONS VS REAL SCORES
```{r}
PREDICTNNTESTING = compute(NN, NNTESTING[,c(1:8)])
PREDICTNNTESTING = (PREDICTNNTESTING$net.result * (max(DF2$score) - min(DF2$score))) + min(DF2$score)

plot(TESTDFNN$score, PREDICTNNTESTING, col='blue', pch=10, ylim = range(0:100), xlim = range(0:100), ylab = "Predicted Score", xlab = "Real Score")


abline(0,1)
```

#FINDING RMSE
```{r}
NN_RMSE = (sum((TESTDFNN$score - PREDICTNNTESTING)^2) / nrow(TESTDFNN)) ^ 0.5
NN_RMSE
```

```{r}
# Normalized RMSE 
#NRMSENN = RMSE / (max value - min value)
NRMSENN = 0.766921 / (100.00 - 44.02)
NRMSENN
```

#VALIDATING NEURAL NETWORK VIA NESTING LOOP
```{r}
set.seed(47)
k = 100
NN_RMSE = NULL

List = list( )

for(j in 10:65){
    for (i in 1:k) {
        INDEXNN = sample(1:nrow(DF2),j )

        NNTRAINING = scaled[INDEXNN,]
        NNTESTING = scaled[-INDEXNN,]
        TESTDFNN = DF2[-INDEXNN,]

        NN = neuralnet(score~quality_of_education + alumni_employment+ quality_of_faculty+ publications+ influence+ citations+ broad_impact+ patents, NNTRAINING, hidden = 3, linear.output= T)
        PREDICTNNTESTING = compute(NN,NNTESTING[,c(1:8)])
        PREDICTNNTESTING = (PREDICTNNTESTING$net.result*(max(DF2$score)-min(DF2$score)))+min(DF2$score)

        NN_RMSE [i]<- (sum((TESTDFNN$score - PREDICTNNTESTING)^2)/nrow(TESTDFNN))^0.5
    }
    List[[j]] = NN_RMSE
}

Matrix.RMSE = do.call(cbind, List)
```

#GRAPHING RMSE VIA BOXPLOT
```{r}
boxplot(Matrix.RMSE[,3], ylab = "RMSE", main = "RMSE BoxPlot (length of training set = 65)")
```

#MEASURING AND PLOTTING VARIATION OF RMSE
```{r}
library(matrixStats)

RMSETIME = colMedians(Matrix.RMSE)

X = seq(10,65)

plot (RMSETIME~X, type = "l", xlab = "Training Set Length", ylab = "median RMSE", main = "RMSE Variation w Training Set Length")
```

#LINEAR MODELING FOR ALL VARIABLES
```{r}
LMALLVAR <- lm(score~quality_of_education + alumni_employment+ quality_of_faculty+ publications+ influence+ citations+ broad_impact+ patents, data=DF1)
summary(LMALLVAR)
```


